Query Optimization for Dynamic Imputation

### 2017-05-11 rebuttal.tex

\item - Although it doesn't consider cost or the drop operator, a similar technique appears in Yang et. al. "Lenses: An On-Demand Approach to ETL." A comparison would be nice.

\item - S1: "Furthermore, the results obtained with this approach are similar in quality to those obtained with the traditional imputation approach — between 0 and 24 percent" 24 percent error is pretty bad.

\item - S2: The drop operator appears to be a weaker form of SQL's null-value / 3-valued Boolean semantics. The latter might actually be more useful. For example, consider the case of a predicate (A $>$ 5 OR B $<$ 10). If A $>$ 5, then the condition is independent of B, and the tuple can still be salvaged. If A $\leq$ 5, then the expression evaluates to UNKNOWN and the tuple is dropped anyway. For aggregates there is a greater need, as one might have different aggregates computed over different subsets of the data, but even here it's not clear that this is any less informative than just throwing data away.

  \resp{We agree that it could be interesting to allow drops to be omitted before filters. This would increase the plan space somewhat and might lead to better plans in some cases. For the conjunctive predicates in our experiments, our implementation splits the predicate into multiple filter nodes, so imputations or drops and filters can interleave. This should have the same effect, as the filter and the drop operator behave the same way on a tuple containing null. However, this optimization would also apply to disjunctive predicates, which cannot be split, and is worth definitely investigating.}

@@

  \resp{We have clarified this in the paper.}

\resp{We emphasize that \textit{Loss} is not a loss in the statistical sense or in the tuple loss sense by renaming it to \textit{Penalty}.}

\item - S3.4.3: An extremely nice emergent property from the L-value function for the imputation operator seems to be that the operator will be moved to a position in the plan where it has access to the maximum training data. I am slightly concerned that by focusing entirely on size, this might lead to skew in the training set. For example, consider a one-to-many join \verb/µ(R) |X| S/. Your optimization algorithm would likely wish to move imputation up: \verb/µ(R |X| S)/. Yes, there is more training data, but if the fanout for each tuple is skewed, some tuples in R will become over-represented in the join result. A similar observation holds for selection. For example take your Q8, where the conditions "studentdebtowe $>$ 0 AND schooldegree $\geq$ 0" presumably indicate invalid data that should not be used for training. (is this perhaps why you have so much error for full imputation on Q8?)

  \resp{This is a reasonable concern. Though often an appropriate proxy, optimizing for the maximum training set is flawed when the query requires the construction a large, skewed data set. The problem seems to be particular to queries which generate large intermediate tables. In some cases, as when joining two datasets, the intermediate table is a good target for training because it exposes correlations which would be hidden in the separate tables. In other cases, this intermediate table is large and skewed, which causes problems for imputation. Avoiding this case requires us to estimate the skew in these intermediate tables.}
  \todo{Need a better explanation here.}
\todo{Add an explanation of the coding of schooldegree -- and why it is filter $>=$ 0, which seems unreasonable to an observer.}

@@

    If necessary, it should be possible to sample from the pareto optimal plans during the plan enumeration. The final selected plan may not be pareto optimal, but this is a potential way to control the number of plans.}

\item D7. When defining the cost of impute operator, how do you know a priori the number of iterations for the algorithm (given it has to converge)?

\item D8. Are there impute implementation that are non-blocking? Or do they always have to perform multiple passes over the data?

  \resp{Yes. For example, in time series data, it is reasonable to fill the most recent non-null value forward. This kind of imputation isn't useful for survey data, so we didn't consider it for our experiments. We have also extended the experiments section to consider a simple imputation method which fills missing values with the mean of the column, which only needs one pass over the data.}

### 2017-05-11 rebuttal.tex

  \resp{We have clarified this in the paper.}

\resp{We emphasize that \textit{Loss} is not a loss in the statistical sense or in the tuple loss sense by renaming it to \textit{Penalty}.}

\item - S3.4.3: An extremely nice emergent property from the L-value function for the imputation operator seems to be that the operator will be moved to a position in the plan where it has access to the maximum training data. I am slightly concerned that by focusing entirely on size, this might lead to skew in the training set. For example, consider a one-to-many join \verb/µ(R) |X| S/. Your optimization algorithm would likely wish to move imputation up: \verb/µ(R |X| S)/. Yes, there is more training data, but if the fanout for each tuple is skewed, some tuples in R will become over-represented in the join result. A similar observation holds for selection. For example take your Q8, where the conditions "studentdebtowe $>$ 0 AND schooldegree $\geq$ 0" presumably indicate invalid data that should not be used for training. (is this perhaps why you have so much error for full imputation on Q8?)

\resp{This is a reasonable concern. Though often an appropriate proxy, optimizing for the maximum training set is flawed when the query requires the construction a large, skewed data set. The problem seems to be particular to queries which generate large intermediate tables. In some cases, as when joining two datasets, the intermediate table is a good target for training because it exposes correlations which would be hidden in the separate tables. In other cases, this intermediate table is large and skewed, which causes problems for imputation. Avoiding this case requires us to estimate the skew in these intermediate tables.

To clarify the point about Q8, we've removed the predicate condition $schooldegree \geq 0$, this was in fact a trivial satisfied predicate, given the way our dataset was set up.}
  \todo{Need a better explanation here.}

### 2017-05-11 paper/main.tex
added a section of implementation components
clean up algorithms
paper: move implementation section

\documentclass{vldb}
\usepackage{balance}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage[activate={true,nocompatibility},final,kerning=true,spacing=true,factor=1100,stretch=10,shrink=10]{microtype}
\usepackage[subtle]{savetrees}
\usepackage{amsfonts,amssymb,amsmath}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[style=numeric,firstinits=true]{biblatex}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{graphicx} % includegraphics
\usepackage{listings} % code samples with syntax hl
\usepackage{stmaryrd}
\usepackage{qtree} % for simple tree diagrams
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xcolor}
\usepackage{hyperref} % load after other packages
\usepackage{tikz}
\usetikzlibrary{positioning}
\usepackage{cleveref}
\usepackage{xspace}
\usepackage[all]{nowidow}

@@

\begin{abstract}
  Missing values are common in data analysis and present a usability challenge.
  Users are forced to pick between removing tuples with missing values or creating a cleaned version of their data by applying a relatively expensive imputation strategy.
  Our system, \ProjectName{}, incorporates imputation into a cost-based query optimizer, performing necessary imputations on-the-fly for each query.
  This allows users to immediately explore their data, while the system picks the optimal placement of imputation operations.
  % This placement is based on the specific query issued and a user-provided parameter that trades off data quality and runtime performance.
  % We incorporate the imputation operators into the relational algebra, allowing for traditional optimizations to take place.
  We evaluate this approach on three real-world survey-based datasets.
  % from the \textit{Center for Disease Control and Prevention} and \textit{freeCodeCamp}.
  Our experiments show that our query plans execute between \lowxalphazero{} and \highxalphazeroexacs{} times faster than first imputing the base tables.
  Furthermore, we show that the query results from on-the-fly imputation differ from the
  traditional base-table imputation approach by \lowsmapealphazero{}--\highsmapealphaoneexacs{}\%, in one error measure.
  Finally, we show that while dropping tuples with missing values that fail query constraints discards 3--88\% of the data, on-the-fly imputation loses only 0--10\%.
\end{abstract}

\input{introduction}
\input{examples}
\input{algorithm}
\input{experiments}
\input{relatedwork}
\input{conclusion}
\bibstyle{abbrv}
\balance
\printbibliography

>>>>

\documentclass{vldb}
\usepackage{balance}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage[activate={true,nocompatibility},final,kerning=true,spacing=true,factor=1100,stretch=10,shrink=10]{microtype}
\usepackage[subtle]{savetrees}
\usepackage{amsfonts,amssymb,amsmath}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[style=numeric,firstinits=true]{biblatex}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{graphicx} % includegraphics
\usepackage{listings} % code samples with syntax hl
\usepackage{stmaryrd}
\usepackage{qtree} % for simple tree diagrams
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usepackage{xspace}
\usepackage[all]{nowidow}
\usepackage{enumitem}
\usepackage{hyperref} % load after other packages
\usepackage{cleveref}

@@

\begin{abstract}
  Missing values are common in data analysis and present a usability challenge.
  Users are forced to pick between removing tuples with missing values or creating a cleaned version of their data by applying a relatively expensive imputation strategy.
  Our system, \ProjectName{}, incorporates imputation into a cost-based query optimizer, performing necessary imputations on-the-fly for each query.
  This allows users to immediately explore their data, while the system picks the optimal placement of imputation operations.
  % This placement is based on the specific query issued and a user-provided parameter that trades off data quality and runtime performance.
  % We incorporate the imputation operators into the relational algebra, allowing for traditional optimizations to take place.
  We evaluate this approach on three real-world survey-based datasets.
  % from the \textit{Center for Disease Control and Prevention} and \textit{freeCodeCamp}.
  Our experiments show that our query plans execute between \todo{update: \lowxalphazero{} and \highxalphazeroexacs{}} times faster than first imputing the base tables.
  Furthermore, we show that the query results from on-the-fly imputation differ from the
  traditional base-table imputation approach by \lowsmapealphazero{}--\highsmapealphaoneexacs{}\%, in one error measure.
  Finally, we show that while dropping tuples with missing values that fail query constraints \todo{update: discards 3--88\% of the data, on-the-fly imputation loses only 0--10\%}.
\end{abstract}

\input{introduction}
\input{examples}
\input{algorithm}
\input{implementation}
\input{experiments}
\input{relatedwork}
\input{conclusion}
\bibstyle{abbrv}
\balance
\printbibliography
